---
title: "Employment"
description: |
  The COVID-19 pandemic has forced millions of workers out of their job, sending the unemployment rate to a level the country last seen in 2011. The formal sector took a harder hit than the informal sector.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
preview: fig/ier_unemployment-rate_void_plot.png    
---


```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse)
library(lubridate)
library(httr)
library(jsonlite)
library(htmltools)
library(xaringanExtra)
library(crosstalk)
library(govdown)
library(mime)

# bps api
## key
if (exists("BPS_KEY") == F) {
  BPS_KEY <- Sys.getenv("BPS_KEY")
}

## base url
if (exists("base_url") == F) {
  base_url <- "https://webapi.bps.go.id/v1/api/list"
}

# use panelset
xaringanExtra::use_panelset()

# description functions
## rise or fall
ier_compare_rise <- function(x) {
  if (x > 0) {
    "rose"
  } else {
    "fell"
  }
}

## up or down
ier_compare_up <- function(x) {
  if (x > 0 ) {
    "up"
  } else {
    "down"
  }
}

# embed file
## unemployment rate
embed_unemp_rate <- xfun::embed_file(
  "data/ier_unemployment-rate_cleaned.csv",
  name = "ier_unemployment-rate_cleaned.csv",
  text = "csv"
)

## number of unemployed
embed_unemp_ppl <- xfun::embed_file(
  "data/ier_unemployed_cleaned.csv",
  name = "ier_unemployed_cleaned.csv",
  text = "csv"
)

## unemployment rate by province
embed_unemp_prov <- xfun::embed_file(
  "data/ier_unemployment-province_cleaned.csv",
  name = "ier_unemployment-province_cleaned.csv",
  text = "csv"
)

## employment by status
embed_emp_status <- xfun::embed_file(
  "data/ier_employment-status_cleaned.csv",
  name = "ier_employment-status_cleaned.csv",
  text = "csv"
)

## employment by sector
embed_emp_sec <- xfun::embed_file(
  "data/ier_employment-sector_cleaned.csv",
  name = "ier_employment-sector_cleaned.csv",
  text = "csv"
)

## pandemic-hit workers
embed_emp_phw <- xfun::embed_file(
  "data/ier_pandemic-hit-workers_cleaned.csv",
  name = "ier_pandemic-hit-workers_cleaned.csv",
  text = "csv"
) 

# embed chart
## unemployment rate
embed_chart_unemp_rate <- xfun::embed_file(
  "fig/ier_unemployment-rate_plot.png",
  name = "ier_unemployment-rate_plot.png",
  text = "chart"
)

## number of unemployed
embed_chart_unemp_ppl <- xfun::embed_file(
  "fig/ier_unemployed_plot.png",
  name = "ier_unemployed_plot.png",
  text = "chart"
)

## unemployment rate by province
embed_chart_unemp_prov <- xfun::embed_file(
  "fig/ier_unemployment-province_plot.png",
  name = "ier_unemployment-province_plot.png",
  text = "chart"
)
```


### Unemployment

::::: {.panelset}

::: {.panel}

#### Rate {.panel-name}

(percent)

```{r plot: unemployment rate}
source("script/emp_unemployment_script.R", local = knitr::knit_global())

div(plot_unemp_rate)
```

:::

::: {.panel}

#### Workers {.panel-name}

(millions)

```{r plot: number of unemployed}
source("script/emp_unemployed_script.R", local = knitr::knit_global())

div(plot_unemp_ppl)
```

:::

:::::

```{r description: unemployment}
# latest unemployment rate data
unemp_rate_latest <- unemp_trf %>% 
  dplyr::filter(date == last(date))

# latest number of unemployed data
unemp_ppl_latest <- unemp_ppl_trf %>% 
  dplyr::filter(date == last(date))
```


The number of unemployed workers `r ier_compare_rise(unemp_ppl_latest$change_yoy)` by **`r unemp_ppl_latest$change_yoy`** million to **`r unemp_ppl_latest$unemployed_ppl`** million in `r format(unemp_ppl_latest$date, "%B %Y")` from a year earlier. As a result, the unemployment rate was `r ier_compare_up(unemp_rate_latest$ppt_change_yoy)` by **`r unemp_rate_latest$ppt_change_yoy`** percentage points to **`r unemp_rate_latest$unemployment_rate`** percent over the same period. The unemployment rate in August 2020 was the highest in a decade.

<aside>

See also [inflation](inflation.html)

</aside>

```{r documentation: unemployemnt}
# date of last update
## request data
pressrelease_req <- GET(
  base_url,
  query = list(
    model = "pressrelease",
    lang = "ind",
    domain = "0000",
    keyword = "pengangguran",
    key = BPS_KEY
  )
)

## parse response
pressrelease_resp <- content(pressrelease_req, "text") %>% 
  fromJSON()

## date of last update
source_last_update <- as_tibble(pressrelease_resp$data[[2]]) %>% 
  dplyr::filter(subj_id == 6) %>% 
  mutate(rl_date = ymd(rl_date)) %>% 
  arrange(desc(rl_date)) %>% 
  dplyr::filter(rl_date == first(rl_date)) %>% 
  select(rl_date) %>% 
  deframe() %>% 
  format("%B %d, %Y")
```

<details>
<summary>Documentation</summary>
**Description**: the number of people aged 15 or above who do not have a job and are looking for a job, preparing a business or not looking for a job because they feel it is impossible to get it, as well as those who already have a job but have not started working. The unemployment rate represents the number of unemployed workers relative to the labor force. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [Statistics Indonesia (BPS)](https://www.bps.go.id/indicator/6/529/1/penduduk-berumur-15-tahun-ke-atas-menurut-jenis-kegiatan.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: unemployment rate (`r embed_unemp_rate`; `r embed_chart_unemp_rate`); number of unemployed workers (`r embed_unemp_ppl`; `r embed_chart_unemp_ppl`)  
</details>


### Unemployment rate by province

(percent)

```{r script: unemployment rate by province}
source("script/emp_unemployment_province_script.R", local = knitr::knit_global())

div(plot_unemp_prov)
```

```{r description: unemployment rate by province}
# latest unemployment rate by province data
unemp_prov_latest <- unemp_prov_trf %>% 
  dplyr::filter(date == unemp_prov_date_latest[2])

# province with highest, lowest increase in unemployment rate
unemp_prov_latest_minmax <- unemp_prov_latest %>% 
  dplyr::filter(
    ppt_change_yoy %in% c(
      min(unemp_prov_latest$ppt_change_yoy),
      max(unemp_prov_latest$ppt_change_yoy)
    )
  ) %>% 
  arrange(desc(ppt_change_yoy))

```

The highest increase in the unemployment rate was seen in `r unemp_prov_latest_minmax$province[1]` at **`r unemp_prov_latest_minmax$ppt_change_yoy[1]`** percentage points, which brought the share of unemployed workers in the province to **`r unemp_prov_latest_minmax$unemployment_rate[1]`** percent in `r format(unemp_prov_latest_minmax$date[1], "%B %Y")`.

<aside>

See also [poverty by province](poverty.html#poverty-by-province)

</aside>

`r last(unemp_prov_latest_minmax$province)` recorded the smallest increase in the unemployment rate at **`r last(unemp_prov_latest_minmax$ppt_change_yoy)`** percentage points yoy to **`r last(unemp_prov_latest_minmax$unemployment_rate)`** percent in August 2020.

<details>
<summary>Documentation</summary>
**Description**: the number of unemployed workers relative to the labor force, broken down by province. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/indicator/6/543/1/tingkat-pengangguran-terbuka-menurut-provinsi.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_unemp_prov`; `r embed_chart_unemp_prov`  
</details>


### Employment by status

```{r table: employment by status, layout="l-body-outset"}
source("script/emp_employment_status_script.R", local = knitr::knit_global())

reactable_emp_status
```

```{r description: employment by status}
# latest employment by status data
emp_status_latest <- emp_status_trf %>% 
  dplyr::filter(date == emp_status_date_latest[2])

# status with steepest decline in distribution
emp_status_min <- emp_status_latest %>% 
  dplyr::filter(prop_chg_yoy == min(emp_status_latest$prop_chg_yoy))
```

The share of `r str_to_lower(emp_status_min$status)` in the labor force showed the steepest annual decline at **`r emp_status_min$prop_chg_yoy`** percentage points to **`r emp_status_min$prop`** percent in `r format(emp_status_min$date, "%B %Y")`. The number of `r str_to_lower(emp_status_min$status)` was down by **`r emp_status_min$workers_pct_chg_yoy`** percent yoy to **`r emp_status_min$workers`** million.

<details>
<summary>Documentation</summary>
**Description**: the number and distribution of workers, broken down by employment status. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/statictable/2009/04/16/971/penduduk-15-tahun-ke-atas-menurut-status-pekerjaan-utama-1986---2020.html){target="_blank"}  
**Last updated**: `r format(emp_status_last_update, "%B %d, %Y")`  
**Download**: `r embed_emp_status`  
</details>


### Employment by sector

```{r script: employment by sector}
source("script/emp_employment_sec_script.R", local = knitr::knit_global())

# filter
unstrap(filter_emp_sec)
```

```{r table: employment by sector, layout="l-body-outset"}
reactable_emp_sec
```

```{r description: employment by sector}
# latest employment by sector data
emp_sec_latest <- emp_sec_trf %>% 
  dplyr::filter(date == emp_sec_date_latest[2])

# sector with steepest decline, highest increase
emp_sec_latest_minmax <- emp_sec_latest %>% 
  dplyr::filter(
    prop_chg_yoy %in% c(
      min(emp_sec_latest$prop_chg_yoy),
      max(emp_sec_latest$prop_chg_yoy)
    )
  ) %>% 
  arrange(prop_chg_yoy)
```

The steepest annual decline in the share of workers was recorded at **`r emp_sec_latest_minmax$prop_chg_yoy[1]`** percentage points in the `r str_to_lower(emp_sec_latest_minmax$sector[1])` sector, which brought the share to **`r emp_sec_latest_minmax$prop[1]`** percent in `r format(emp_sec_latest_minmax$date[1], "%B %Y")`. The number of workers in the sector was down by **`r emp_sec_latest_minmax$workers_pct_chg_yoy[1]`** percent yoy to **`r emp_sec_latest_minmax$workers[1]`** million over the period.

<aside>

See also [economic growth by sector](gdp.html#by-sector)

</aside>

Meanwhile, the `r str_to_lower(last(emp_sec_latest_minmax$sector))` sector recorded a **`r last(emp_sec_latest_minmax$workers_pct_chg_yoy)`** percent yoy increase in the number of workers to **`r last(emp_sec_latest_minmax$workers)`** million, leading to an increase of **`r last(emp_sec_latest_minmax$prop_chg_yoy)`** percentage points yoy in the share of workers to **`r last(emp_sec_latest_minmax$prop)`** percent.

<details>
<summary>Documentation</summary>
**Description**: the number and distribution of workers, broken down by sector. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/statictable/2009/04/16/970/penduduk-15-tahun-ke-atas-yang-bekerja-menurut-lapangan-pekerjaan-utama-1986---2020.html){target="_blank"}  
**Last updated**: `r format(emp_sec_last_update, "%B %d, %Y")`  
**Download**: `r embed_emp_sec`  
</details>


### Number of workers affected by COVID-19

(millions)

```{r script: overall affected workers}
source("script/emp_pandemic-hit_script.R", local = knitr::knit_global())

div(plot_phw_total)
```

*Tips: click the legend item to select or deselect a trace.*

The number of workers affected by the pandemic was `r ier_compare_up(phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "total"])` by **`r phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "total"]`** percent to **`r phw_trf_tot_dis$workers[phw_trf_tot_dis$category == "total"]`** million in February from August. They accounted for **`r phw_share_tidy$share_of_working_age_pop[phw_share_tidy$date == ymd("2021-02-01") & phw_share_tidy$category == "total"]`** percent of the working age population. The majority of them faced reduced working hours.

<details>
<summary>Documentation</summary>
**Description**: people who have been laid off or furloughed because of the pandemic in the period starting from February 2020, broken down by status. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/pressrelease/2021/05/05/1815/februari-2021--tingkat-pengangguran-terbuka--tpt--sebesar-6-26-persen.html)  
**Last updated**: `r source_last_update`  
**Download**: `r embed_emp_phw`
</details>


### Number of pandemic-hit workers by gender

```{r script: affected workers by gender, layout="l-body-outset"}
reactable_phw_sex
```

<div style="color: #999; font-size: 12px;">
<sup>*</sup>Compared with August 2020
</div>

Between August and February, the total number of pandemic-hit male workers `r ier_compare_rise(phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "men"])` by **`r phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "men"]`** percent to **`r phw_trf_tot_dis$workers[phw_trf_tot_dis$category == "men"]`** million and female workers by **`r phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "women"]`** percent to **`r phw_trf_tot_dis$workers[phw_trf_tot_dis$category == "women"]`** million.

<details>
<summary>Documentation</summary>
**Description**: people who have been laid off or furloughed because of the pandemic in the period starting from February 2020, broken down by gender. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/pressrelease/2021/05/05/1815/februari-2021--tingkat-pengangguran-terbuka--tpt--sebesar-6-26-persen.html)  
**Last updated**: `r source_last_update`  
**Download**: `r embed_emp_phw`
</details>


### Number of pandemic-hit workers by area

```{r script: affected workers by area, layout="l-body-outset"}
reactable_phw_area
```

<div style="color: #999; font-size: 12px;">
<sup>*</sup>Compared with August 2020
</div>

The total number of pandemic-hit workers in urban areas `r ier_compare_rise(phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "urban"])` by **`r phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "urban"]`** percent to **`r phw_trf_tot_dis$workers[phw_trf_tot_dis$category == "urban"]`** million  and in rural areas by **`r phw_trf_tot_dis$pct_chg[phw_trf_tot_dis$category == "rural"]`** percent **`r phw_trf_tot_dis$workers[phw_trf_tot_dis$category == "rural"]`** million in February from August.

<details>
<summary>Documentation</summary>
**Description**: people who have been laid off or furloughed because of the pandemic in the period starting from February 2020, broken down by area. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every May and November  
**Source**: [BPS](https://www.bps.go.id/pressrelease/2021/05/05/1815/februari-2021--tingkat-pengangguran-terbuka--tpt--sebesar-6-26-persen.html)  
**Last updated**: `r source_last_update`  
**Download**: `r embed_emp_phw`
</details>


# Note {.appendix}

The data is based on the National Labor Force Survey (Sakernas) from BPS, which takes a sample of 75,000 households in the February round and 300,000 households in the August round. Read more about the data [here](https://bps.go.id/subject/6/tenaga-kerja.html#subjekViewTab1){target="_blank"}.

The International Labor Organization (ILO) has [replaced](https://www.ilo.org/wcmsp5/groups/public/---dgreports/---stat/documents/publication/wcms_422401.pdf){target="_blank"} unpaid family workers with contributing family workers, namely “those workers who hold ‘self-employment jobs’ as own-account workers [...] in a market-oriented establishment operated by a related person living in the same household.”

BPS started in August 2020 adding the data on workers affected by COVID-19, namely those who have been laid off or furloughed in the period starting from February 2020. Aside from workers getting unemployed and facing reduced working hours, the agency also recorded other pandemic impacts on the labor force:

- Not in the workforce: people aged 15 or above who still go to school, take care of their household or have other non-personal activities. Working age population who have stopped working and are out of the workforce in the period starting from February 2020 belong to this group.

- Furloughed: workers who have not been working in the past week for various reasons. People who have jobs but have not been working temporarily in the period starting from February 2020 belong to this group.