---
title: "Economic growth"
description: |
  Indonesia’s economy plunged into recession in the third quarter of 2020, the country’s first recession since the Asian financial crisis. The economy has shown signs of recovery since then, but the pandemic emergency curbs have forced the government and analysts to slash growth forecast for 2021.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
preview: fig/ier_gdp-growth_void_plot.png
---

```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(htmltools)
library(govdown)
library(mime)

# functions to describe data
## period
ier_period <- function(x) {
  if (str_detect(x, "-01-01")) {
    "January-March period"
  } else if (str_detect(x, "-04-01")) {
    "April-June period"
  } else if (str_detect(x, "-07-01")) {
    "July-September period"
  } else {
    "October-December period"
  }
}

## up or down
## grow or contract
ier_compare_grow <- function(x) {
  if (x > 0 ) {
    "grew"
  } else {
    "contracted"
  }
}
  
# embed file
## econ growth by expenditure
embed_growth_exp <- xfun::embed_file(
  "data/ier_gdp-growth-exp_cleaned.csv",
  name = "ier_gdp-growth-exp_cleaned.csv",
  text = "csv"
)

## contribution to growth
embed_growth_ctg <- xfun::embed_file(
  "data/ier_gdp-ctg_cleaned.csv",
  name = "ier_gdp-ctg_cleaned.csv",
  text = "csv"
)

## econ growth by sector
embed_growth_sec <- xfun::embed_file(
  "data/ier_gdp-growth-sector_cleaned.csv",
  name = "ier_gdp-growth-sector_cleaned.csv",
  text = "csv"
)

# embed chart
## growth
embed_chart_growth <- xfun::embed_file(
  "fig/ier_gdp-growth_plot.png",
  name = "ier_gdp-growth_plot.png",
  text = "chart"
)

## contribution to growth
embed_chart_ctg <- xfun::embed_file(
  "fig/ier_gdp-ctg_plot.png",
  name = "ier_gdp-ctg_plot.png",
  text = "chart"
)
```

### Overall

Gross domestic product (GDP) (percent change from a year earlier)

```{r plot: overall econ growth}
source("script/gdp_growth_script.R", local = knitr::knit_global())

div(plot_growth)
```

```{r description: overall econ growth}
# subset two latest values
growth_latest <- growth_tidy$pct_change_yoy %>% 
  tail(2)
```

The country's GDP `r ier_compare_grow(last(growth_tidy$pct_change_yoy))` by **`r last(growth_tidy$pct_change_yoy)`** percent year-on-year (yoy) in the `r ier_period(last(growth_tidy$date))`.

<aside>

See also [community mobility](mobility.html)

</aside>

The pandemic led the economy to contract by **5.32** percent yoy in the second quarter of 2020 and pushed it to its first recession in two decades as it contracted by **3.49** percent yoy in the third quarter of last year.

```{r documentation: gdp}
# date of last update
## request data
pressrelease_req <- GET(
  base_url,
  query = list(
    model = "pressrelease",
    lang = "ind",
    domain = "0000",
    keyword = "ekonomi",
    key = BPS_KEY
  )
)

## parse response
pressrelease_resp <- content(pressrelease_req, "text") %>% 
  fromJSON()

## date of last update
source_last_update <- as_tibble(pressrelease_resp$data[[2]]) %>% 
  dplyr::filter(subj_id == 11) %>% 
  mutate(rl_date = ymd(rl_date)) %>% 
  arrange(desc(rl_date)) %>% 
  dplyr::filter(rl_date == first(rl_date)) %>% 
  select(rl_date) %>% 
  deframe() %>% 
  format("%B %d, %Y")
```

<details>
<summary>Documentation and download options</summary>
**Description**: the value of final goods and services produced by all residents in a given quarter in constant 2010 prices, as measured in terms of percent change from the same quarter a year earlier. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Data and update frequency**: quarterly  
**Source**: [Statistics Indonesia (BPS)](https://bps.go.id/indicator/169/108/1/-seri-2010-laju-pertumbuhan-pdb-menurut-pengeluaran.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_growth_exp`; `r embed_chart_growth`  
</details>


#### By expenditure

GDP (percent change from a year earlier)

```{r script: econ growth by expenditure}
source("script/gdp_growth_exp_script.R", local = knitr::knit_global())

# filter
filter_growth_exp <- filter_select(
  "exp",
  "Select expenditure...",
  growth_exp_shared,
  ~label_eng
)

## button
unstrap(filter_growth_exp)
```

```{r table: econ growth by expenditure, layout="l-body-outset"}
reactable_growth_exp
```

```{r description: econ growth by expenditure}
# latest household consumption values
growth_exp_hh_latest <- growth_exp_tidy %>% 
  dplyr::filter(label_eng == "Household consumption") %>% 
  tail(2)
```

<div style="color: #999; font-size: 12px;">
<sup>*</sup>Non-profit institutions serving households (NPISHs)
</div>

The household consumption, the largest contributor to GDP by expenditure, `r ier_compare_grow(growth_exp_hh_latest$pct_change_yoy[2])` by **`r growth_exp_hh_latest$pct_change_yoy[2]`** percent yoy in the `r ier_period(last(growth_exp_tidy$date))`.

<details>
<summary>Documentation and download options</summary>
**Description**: the value of final expenditure of goods and services at purchasers’ prices minus the value of imports of goods and services in a given quarter in constant 2010 prices, as measured in terms of percent change from the same quarter a year earlier. See [appendix](#note) for more details.  
**Table**: Dzulfiqar Fathur Rahman  
**Data and update frequency**: quarterly  
**Source**: [BPS](https://bps.go.id/indicator/169/108/1/-seri-2010-laju-pertumbuhan-pdb-menurut-pengeluaran.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_growth_exp`  
</details>


#### By sector

GDP (percent change from a year earlier)

```{r script: econ growth by sector}
source("script/gdp_growth_sec_script.R", local = knitr::knit_global())

# filter
filter_growth_sec <- filter_select(
  "exp",
  "Select sector...",
  growth_sec_shared,
  ~key_sector
)

## button
unstrap(filter_growth_sec)
```

```{r table: econ growth by sector, layout="l-body-outset"}
# tabel
reactable_growth_sec
```

```{r description: econ growth by sector}
# latest manufacturing value
manufacturing <- growth_sec_tidy %>% 
  dplyr::filter(key_sector == "Manufacturing") %>%
  tail(2) %>% 
  select(ncol(.)) %>% 
  deframe()
```

The manufacturing sector, the largest contributor to GDP by sector, `r ier_compare_grow(manufacturing[2])` by **`r manufacturing[2]`** percent yoy in the `r ier_period(last(growth_sec_tidy$date))`.

<aside>

See also [employment by sector](employment.html#by-sector)

</aside>

<details>
<summary>Documentation and download options</summary>
**Description**: the total value added generated by all resident production units broken down by industrial classification of economic activities in a given quarter in constant 2010 prices, as measured in terms of percent change from the same quarter a year earlier. See [appendix](#note) for more details.  
**Table**: Dzulfiqar Fathur Rahman  
**Data and update frequency**: quarterly  
**Source**: [BPS](https://bps.go.id/indicator/11/104/1/-seri-2010-laju-pertumbuhan-pdb-seri-2010.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_growth_sec`  
</details>

### Contribution to growth

By expenditure (percentage point)

```{r plot: contribution to growth}
source("script/gdp_ctg_exp_script.R", local = knitr::knit_global())
div(plot_ctg)
```

*Tip: click the legend item(s) to select or remove data*

```{r description: contribution to growth}
# latest values
latest_ctg_pc <- last(ctg_exp_trf$`Private consumption`)
latest_ctg_inv <- last(ctg_exp_trf$Investment)
latest_ctg_nx <- last(ctg_exp_trf$`Net export`)
latest_ctg_govt <- last(ctg_exp_trf$`Government consumption`)
```

Private consumption contributed **`r latest_ctg_pc`** percentage points and investment **`r latest_ctg_inv`** percentage points to annual GDP growth in the `r ier_period(last(ctg_exp_trf$date))`. 

Net export and government consumption offset the contractions in other expenditure components by contributing **`r latest_ctg_nx`** and **`r latest_ctg_govt`** percentage points, respectively.

<details>
<summary>Documentation and download options</summary>
**Description**: the contribution to growth is a component’s growth weighted by its share in the GDP at constant 2010 prices in the previous quarter. Private consumption consists of household consumption and NPISHs consumption. Investment consists of gross fixed capital formation and changes in inventories.  
**Chart**: Dzulfiqar Fathur Rahman  
**Data and update frequency**: quarterly  
**Source**: [BPS](https://bps.go.id/indicator/169/1956/1/-seri-2010-pdb-triwulanan-atas-dasar-harga-konstan-menurut-pengeluaran.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_growth_ctg`; `r embed_chart_ctg`  
</details>


# Note {.appendix}

BPS publishes the data dictionary and the method it uses to calculate Indonesia’s GDP with the expenditure, production and income approaches on its website. Read more about the data [here](https://bps.go.id/subject/169/produk-domestik-bruto--pengeluaran-.html#subjekViewTab1) and [here](https://bps.go.id/subject/11/produk-domestik-bruto--lapangan-usaha-.html#subjekViewTab2).