---
title: "Inflation"
description: |
  Consumer price inflation has stay muted and below Bank Indonesia’s target since the middle of 2020, around two months into the COVID-19 pandemic.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
preview: fig/ier_inflation-annual_void_plot.png
---

```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(htmltools)
library(xaringanExtra)
library(crosstalk)
library(govdown)
library(mime)

# use panelset
xaringanExtra::use_panelset()

# functions to describe data
## inflation or deflation
ier_compare_inf <- function(x) {
  if (x > 0) {
    "inflation"
  } else {
    "deflation"
  }
}

## lower or higher
ier_compare_high <- function(x, y) {
  if (x > y) {
    "higher"
  } else {
    "lower"
  }
}

# embed file
## overall inflation
embed_inf_overall <- xfun::embed_file(
  "data/ier_inflation-overall_cleaned.csv",
  name = "ier_inflation-overall_cleaned.csv",
  text = "csv"
)

## inflation by component
embed_inf_comp <- xfun::embed_file(
  "data/ier_inflation-component_cleaned.csv",
  name = "ier_inflation-component_cleaned.csv",
  text = "csv"
)

## inflation by expenditure
embed_inf_exp <- xfun::embed_file(
  "data/ier_inflation-expenditure_cleaned.csv",
  name = "ier_inflation-expenditure_cleaned.csv",
  text = "csv"
)

# embed chart
## monthly overall inflation
embed_chart_inf_overall_mom <- xfun::embed_file(
  "fig/ier_inflation-monthly_plot.png",
  name = "ier_inflation-monthly_plot.png",
  text = "chart"
)

## annual overall inflation
embed_chart_inf_overall_yoy <- xfun::embed_file(
  "fig/ier_inflation-annual_plot.png",
  name = "ier_inflation-annual_plot.png",
  text = "chart"
)

## inflation by component
embed_chart_inf_comp <- xfun::embed_file(
  "fig/ier_inflation-component_plot.png",
  name = "ier_inflation-component_plot.png",
  text = "chart"
)
```

### Overall

(percent)

::::: {.panelset}

::: {.panel}

#### Monthly {.panel-name}

```{r plot: monthly overall inflation}
source("script/inflation_script.R", local = knitr::knit_global())
div(plot_inf_mom)
```

:::

::: {.panel}

#### Annual {.panel-name}

```{r plot: annual overall inflation}
div(plot_inf_yoy)
```

:::

:::::

```{r description: overall inflation}
# latest values
## monthly inflation
inf_mom_latest <- inf_mom_yoy %>%
  select(-rate_yoy) %>% 
  tail(2) %>% 
  select(-c("mo", "yr"))

## annual inflation
inf_yoy_latest <- inf_mom_yoy %>%
  select(-rate_mom) %>% 
  dplyr::filter(!is.na(rate_yoy)) %>% 
  tail(2) %>% 
  select(-c("mo", "yr"))

# date of last update
## dynamic
### request data
pressrelease_req <- GET(
  base_url_dynamic,
  query = list(
    model = "pressrelease",
    lang = "ind",
    domain = "0000",
    keyword = "inflasi",
    key = BPS_KEY
  )
)

### parse response
pressrelease_resp <- content(pressrelease_req, "text") %>% 
  fromJSON()

### date of last update
source_last_update_dynamic <- as_tibble(pressrelease_resp$data[[2]]) %>% 
  dplyr::filter(subj_id == 3) %>% 
  mutate(rl_date = ymd(rl_date)) %>% 
  arrange(desc(rl_date)) %>% 
  dplyr::filter(rl_date == first(rl_date)) %>% 
  select(rl_date) %>% 
  deframe() %>% 
  format("%B %d, %Y")

## static
source_last_update_static <- inf_yoy_parsed$data %>% 
  as_tibble() %>% 
  select(updt_date) %>% 
  ymd_hms() %>% 
  format("%B %d, %Y")
```

The Consumer Price Index (CPI) posted a monthly `r ier_compare_inf(inf_mom_latest$rate_mom[2])` of **`r inf_mom_latest$rate_mom[2]`** percent in `r format(inf_mom_latest$date[2], "%B")`. On an annual basis, the `r ier_compare_inf(inf_yoy_latest$rate_yoy[2])` rate stood at **`r inf_yoy_latest$rate_yoy[2]`** percent.

<aside>

See also [unemployment](employment.html#unemployment)

</aside>

In 2020, the pandemic led to a record-low full-year annual inflation rate at **1.68** percent.

<details>
<summary>Documentation and download options</summary>
**Description**:  the rise in consumer prices as measured in terms of the percent change on the CPI either from a month earlier or a year earlier. Since 2020, the base year for the index is 2018. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Source**: Statistics Indonesia (BPS) ([monthly](https://www.bps.go.id/statictable/2014/09/26/913/tingkat-inflasi-bulanan-gabungan-90-kota-sup-1-sup-2018-100-.html){target="_blank"}; [annual](https://www.bps.go.id/statictable/2014/09/26/915/tingkat-inflasi-tahun-ke-tahun-gabungan-90-kota-sup-1-sup-2018-100-.html){target="_blank"})  
**Data and update frequency**: monthly  
**Last updated**: monthly inflation (`r source_last_update_dynamic`); annual inflation (`r source_last_update_static`)  
**Download**: `r embed_inf_overall`; monthly inflation (`r embed_chart_inf_overall_mom`); annual inflation(`r embed_chart_inf_overall_yoy`)  
</details>


#### By component

CPI (percent change from a month earlier)

```{r plot: inflation by component}
source("script/inflation_comp_script.R", local = knitr::knit_global())
div(sm_inf_comp)
```

```{r description: inflation by component}
# latest values
inf_comp_tidy_latest <- inf_comp_tidy %>% 
  select(-c("mo", "yr")) %>% 
  tail(2)
```

The monthly core `r ier_compare_inf(inf_comp_tidy_latest$core[2])` rate was recorded at **`r inf_comp_tidy_latest$core[2]`** percent in `r format(inf_comp_tidy_latest$date[2], "%B")`. 

For administered prices, the monthly `r ier_compare_inf(inf_comp_tidy_latest$adm_prices[2])` rate stood at **`r inf_comp_tidy_latest$adm_prices[2]`** percent. For volatile prices, the monthly `r ier_compare_inf(inf_comp_tidy_latest$volatile_prices[2])` rate stood at **`r inf_comp_tidy_latest$volatile_prices[2]`** percent.

<details>
<summary>Documentation and download options</summary>
**Description**:  the rise in consumer prices as measured in terms of the monthly percent change on the CPI, broken down by the component. Since 2020, the base year for the index is 2018. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Source**: [BPS](https://www.bps.go.id/statictable/2012/02/02/908/inflasi-umum-inti-harga-yang-diatur-pemerintah-dan-barang-bergejolak-inflasi-indonesia-2009-2021.html){target="_blank"}  
**Data and update frequency**: monthly  
**Last updated**: `r source_last_update_static`  
**Download**: `r embed_inf_comp`; `r embed_chart_inf_comp`  
</details>


#### By expenditure group

CPI (percent change from a month earlier)

```{r script: inflation by expenditure}
source("script/inflation_exp_script.R", local = knitr::knit_global())

# filter
filter_inf_exp <- filter_select(
  "exp",
  "Select expenditure group...",
  inf_exp_wide_shared,
  ~exp_group
)

## button
unstrap(filter_inf_exp)
```

```{r table: inflation by expenditure, layout="l-page"}
reactable_inf_exp
```

```{r description: inflation by expenditure}
# expenditure groups with highest inflation rates
inf_exp_highest <- inf_exp_tidy %>% 
  dplyr::filter(date == last(date)) %>% 
  arrange(desc(rate_mom)) %>% 
  head()
```

The highest monthly inflation rate was seen in the `r str_to_lower(inf_exp_highest$exp_group[1])` at **`r inf_exp_highest$rate_mom[1]`** percent in `r format(inf_exp_highest$date[1], "%B")`. It was followed by `r str_to_lower(inf_exp_highest$exp_group[2])` and `r str_to_lower(inf_exp_highest$exp_group[3])` expenditure groups.

<aside>

See also [retail sales index by category](rsi.html#by-category)

</aside>

<details>
<summary>Documentation and download options</summary>
**Description**:  the rise in consumer prices as measured in terms of the monthly percent change on the CPI, broken down by the expenditure group. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Source**: [BPS](https://www.bps.go.id/statictable/2020/02/04/2083/inflasi-indonesia-menurut-kelompok-pengeluaran-2020-2021.html){target="_blank"}  
**Data and update frequency**: monthly  
**Last updated**: `r source_last_update_static`  
**Download**: `r embed_inf_exp`  
</details>


# Note {.appendix}

BPS builds the CPI based on the 2018 Cost of Living Survey, which covered 90 cities and 853 commodities. Of the commodities, 711 were core components, 101 were volatile prices and the rest were government-administered prices.

Bank Indonesia has set a target of maintaining the annual inflation rate between 2 and 4 percent.

Read more about the data [here](https://bps.go.id/subject/3/inflasi.html#subjekViewTab2){target="_blank"}.