---
title: "Community mobility"
description: |
  Indonesians have been reducing their visit to places like retail and recreation places in response to pandemic restrictions. It remains to be seen whether mobility will return to its normal level since there is a shift in people’s behavior.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
preview: fig/ier_mobility-ntl_void_plot.png
---

```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(htmltools)
library(govdown)
library(mime)

# source url
source_url <- "https://www.google.com/covid19/mobility/"

# embed file
## national mobility
embed_mob_ntl <- xfun::embed_file(
  "data/ier_mobility-ntl_cleaned.csv",
  name = "ier_mobility-ntl_cleaned.csv",
  text = "csv"
)

## regional mobility
embed_mob_reg <- xfun::embed_file(
  "data/ier_mobility-reg_cleaned.csv",
  name = "ier_mobility-reg_cleaned.csv",
  text = "csv"
)

# embed chart
embed_chart_mob_ntl <- xfun::embed_file(
  "fig/ier_mobility-ntl_plot.png",
  name = "ier_mobility-ntl_plot.png",
  text = "chart"
)
```


### National

Number of visitors, by place category  
(percent change from pre-pandemic usual level)

```{r plot: national mobility}
source("script/mobility_ntl_script.R", local = knitr::knit_global())
div(sm_mob_ntl)
```

*Tip: click the legend item(s) to select or remove data*

```{r description: national mobility}
# categories with declining movement trends
mob_ntl_down <- mob_ntl_trf %>% 
  pivot_longer(2:ncol(.), names_to = "category", values_to = "mobility") %>% 
  dplyr::filter(
    date == last(date),
    str_detect(category, "avg"),
    mobility < 0
  )

# turn categories into human-readable format with sentence case
mob_ntl_down$category <- mob_ntl_down$category %>% 
  str_remove("_avg") %>% 
  str_replace_all("_", " and ") %>% 
  str_replace("transit and stations", "transit stations")
```

On `r format(mob_ntl_down$date[1], "%A (%b %d)")`, the number of visitors to `r mob_ntl_down$category[-ncol(mob_ntl_down)]` and `r mob_ntl_down$category[ncol(mob_ntl_down)]` locations was still below the pre-pandemic level on a seven-day moving average basis.

<aside>

See also [retail sales](rsi.html)

</aside>

<details>
<summary>Documentation and download options</summary>
**Description**: the headline number represents a percent change in the number of visitors from the median value of that day of the week over the period between Jan. 3 and Feb. 6, 2020, which reflects the usual level before the pandemic. The data is smoothed to a seven-day moving average. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Data frequency**: daily  
**Update frequency**: once in two or three days  
**Source**: [Google](`r source_url`){target="_blank"}  
**Last updated**: `r format(Sys.Date(), "%B %d, %Y")`  
**Download**: `r embed_mob_ntl`; `r embed_chart_mob_ntl`  
</details>


#### By province

Number of visitors, by place category  
(percent change from pre-pandemic usual level)

```{r script: regional mobility}
source("script/mobility_reg_script.R", local = knitr::knit_global())

# filter by province
unstrap(filter_by_reg)
```

```{js filter: default value for regional mobility}
$(document).ready(function() {
  document.getElementById("reg").getElementsByClassName("selectized")[0].selectize.setValue("Jakarta", false);
});
```

```{r plot: regional mobility}
div(sm_mob_reg)
```

Movement trends are largely similar across provinces, with a few exceptions. 

<details>
<summary>Documentation and download options</summary>
**Description**: the headline number represents a percent change in the number of visitors from the median value of that day of the week over the period between Jan. 3 and Feb. 6, 2020, which reflects the usual level before the pandemic. The data is broken down by province and smoothed to a seven-day moving average. See [appendix](#note) for more details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Data frequency**: daily  
**Update frequency**: once in two or three days  
**Source**: [Google](`r source_url`){target="_blank"}  
**Last updated**: `r format(Sys.Date(), "%B %d, %Y")`  
**Download**: `r embed_mob_reg`  
</details>


# Note {.appendix}

Google constructs the community mobility report based on anonymized data.

While most place categories show a change in the number of visitors, the residential category shows a change in the hours spent in residential places.

Google will intentionally leave gaps in the dataset when the data does not meet the quality threshold and there is not enough data to ensure anonymity.

Google also recommends to not compare countries or regions because of local differences.

The data also does not account for seasonal factors.

Since the headline number represents a change from the baseline value for a particular day of the week, the same number of visitors on two different days of the week to a place category will result in different percent changes.

Therefore, Google advises to avoid comparing day-to-day changes especially between weekends and weekdays.

Read more to understand the data [here](https://support.google.com/covid19-mobility/answer/9825414?hl=en&ref_topic=9822927){target="_blank"}.
