---
title: "Poverty"
description: |
  The COVID-19 pandemic has pushed millions into poverty, undoing around three years worth of progress in lifting people out of it. The rise in the poverty rate was higher in urban areas than rural areas.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
preview: fig/ier_poverty-rate_void_plot.png
---


```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse)
library(lubridate)
library(jsonlite)
library(httr)
library(htmltools)
library(xaringanExtra)
library(crosstalk)
library(govdown)
library(mime)

# bps api
## key
if (exists("BPS_KEY") == F) {
  BPS_KEY <- Sys.getenv("BPS_KEY")
}

## base url
if (exists("base_url") == F) {
  base_url <- "https://webapi.bps.go.id/v1/api/list"
}

# use panelset
xaringanExtra::use_panelset()

# description functions
## rise or fall
ier_compare_rise <- function(x) {
  if (x > 0) {
    "rose"
  } else {
    "fell"
  }
}

## up or down
ier_compare_up <- function(x) {
  if (x > 0 ) {
    "up"
  } else {
    "down"
  }
}

# embed file
## poverty rate
embed_pov_rate <- xfun::embed_file(
  "data/ier_poverty-rate_cleaned.csv",
  name = "ier_poverty-rate_cleaned.csv",
  text = "csv"
)

## number of poor people
embed_poor <- xfun::embed_file(
  "data/ier_poor_cleaned.csv",
  name = "ier_poor_cleaned.csv",
  text = "csv"
)

## poverty by province
embed_pov_prov <- xfun::embed_file(
  "data/ier_poverty-province_cleaned.csv",
  name = "ier_poverty-province_cleaned.csv",
  text = "csv"
)

# embed chart
## poverty rate
embed_chart_pov_rate <- xfun::embed_file(
  "fig/ier_poverty-rate_plot.png",
  name = "ier_poverty-rate_plot.png",
  text = "chart"
)

## number of poor people
embed_chart_poor <- xfun::embed_file(
  "fig/ier_poverty-poor_plot.png",
  name = "ier_poverty-poor_plot.png",
  text = "chart"
)
```


### Poverty rate

::: {.panelset}

::: {.panel}

#### Rate

By area (percent)

```{r plot: poverty rate}
source("script/poverty_script.R", local = knitr::knit_global())

div(plot_pov_rate)
```

:::

::: {.panel}

#### Change

By area (percentage point change from a year earlier)

```{r plot: change in poverty rate}
div(plot_pov_diff)
```

:::

:::

*Tips: click the legend item to select or deselect a trace.*

```{r description: poverty rate}
# latest poverty rate data
pov_latest <- pov_trf %>% 
  dplyr::filter(date == last(date))
```

The overall poverty rate `r ier_compare_rise(pov_latest$diff[3])` by **`r pov_latest$diff[3]`** percentage points year-on-year (yoy) to **`r pov_latest$pov_rate[3]`** percent in `r format(pov_latest$date[1], "%B %Y")`.

<aside>

See also [unemployment](employment.html)

</aside>

In urban areas, the poverty rate `r ier_compare_rise(pov_latest$diff[1])` by **`r pov_latest$diff[1]`** percentage points yoy to **`r pov_latest$pov_rate[1]`** percent. In rural areas, the poverty rate was `r ier_compare_up(pov_latest$diff[2])` by **`r pov_latest$diff[2]`** percentage points yoy to **`r pov_latest$pov_rate[2]`** percent.

In September 2020, the pandemic has pushed the poverty rate back to around the level last seen in the same month in 2017.

```{r documentation: poverty}
# date of last update
## request data
pressrelease_req <- GET(
  base_url,
  query = list(
    model = "pressrelease",
    lang = "ind",
    domain = "0000",
    keyword = "miskin",
    key = BPS_KEY
  )
)

## parse response
pressrelease_resp <- content(pressrelease_req, "text") %>% 
  fromJSON()

## date of last update
source_last_update <- as_tibble(pressrelease_resp$data[[2]]) %>% 
  dplyr::filter(subj_id == 23) %>% 
  mutate(rl_date = ymd(rl_date)) %>% 
  arrange(desc(rl_date)) %>% 
  dplyr::filter(rl_date == first(rl_date)) %>% 
  select(rl_date) %>% 
  deframe() %>% 
  format("%B %d, %Y")
```


<details>
<summary>Documentation</summary>
**Description**: the share of people who cannot afford both food and non-food basic needs as measured by monthly spending. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every February and July  
**Source**: [Statistics Indonesia (BPS)](https://www.bps.go.id/indicator/23/184/1/persentase-penduduk-miskin-menurut-wilayah.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_pov_rate`; `r embed_chart_pov_rate`  
</details>


### Number of poor people

::: {.panelset}

::: {.panel}

#### Total

By area (millions)

```{r plot: number of poor people}
source("script/poverty_poor_script.R", local = knitr::knit_global())

div(plot_poor_total)
```

:::

::: {.panel}

#### Change

By area (millions)

```{r plot: change in the number of poor people}
div(plot_poor_diff)
```

:::

:::

*Tips: click the legend item to select or deselect a trace.*

```{r description: number of poor people}
poor_latest <- poor_trf %>% 
  dplyr::filter(date == last(date))
```

The total number of poor people was `r ier_compare_up(poor_latest$diff[3])` by **`r poor_latest$diff[3]`** million yoy to **`r poor_latest$poor[3]`** million in `r format(poor_latest$date[1], "%B %Y")`. In urban areas, the number `r ier_compare_rise(poor_latest$diff[1])` by **`r poor_latest$diff[1]`** million yoy to **`r poor_latest$poor[1]`** million. In rural areas, it `r ier_compare_rise(poor_latest$diff[2])` by **`r poor_latest$diff[2]`** million yoy to **`r poor_latest$poor[2]`** million.

<details>
<summary>Documentation</summary>
**Description**: the absolute number of people who cannot afford both food and non-food basic needs as measured by monthly spending. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every February and July  
**Source**: [BPS](https://www.bps.go.id/indicator/23/185/1/jumlah-penduduk-miskin-menurut-provinsi.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_poor`; `r embed_chart_poor`  
</details>


### Poverty by province

```{r script: poverty by province}
source("script/poverty_province_script.R", local = knitr::knit_global())

# filter
filter_pov_poor_prov <- filter_select(
  "pov_poor_prov",
  "Select a province...",
  pov_poor_sub_wide_shared,
  ~prov
)

## button
unstrap(filter_pov_poor_prov)
```

```{r table: poverty by province, layout="l-body-outset"}
reactable_pov_poor
```

```{r description: poverty by province}
pov_poor_head <- pov_poor_sub_wide %>% 
  arrange(desc(pov_rate_diff)) %>% 
  head(5)
```

The highest increase in the overall poverty rate was seen in `r pov_poor_head$prov[1]`, where it `r ier_compare_rise(pov_poor_head$pov_rate_diff[1])` by **`r pov_poor_head$pov_rate_diff[1]`** percentage points yoy, which brought the share of the poor population in the province to **`r pov_poor_head$pov_rate_2[1]`** percent in `r format(pov_latest$date[1], "%B %Y")`. It was followed by `r pov_poor_head$prov[-c(1, 5)]` and `r pov_poor_head$prov[5]`.

<aside>

See also [unemployment by province](employment.html#unemployment-rate-by-province)

</aside>

<details>
<summary>Documentation</summary>
**Description**: the absolute number of people who cannot afford both food and non-food basic needs as measured by monthly spending and the poverty rate, broken down by province. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every February and July  
**Source**: [BPS](https://www.bps.go.id/indicator/23/192/1/persentase-penduduk-miskin-menurut-provinsi.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_pov_prov`  
</details>


# Note {.appendix}

The data is based on the National Social and Economic Survey (Susenas) from BPS, which takes a sample of roughly 300,000 households for the March round and 75,000 households for the September round.

BPS uses the basic needs approach, which defines poverty as the inability to afford both food and non-food basic needs as measured by spending.

```{r description: poverty line}
source("script/poverty_line.R", local = knitr::knit_global())

pov_line_latest <- pov_line_trf %>% 
  dplyr::filter(date == last(date))
```

Poor people were those who spent below Rp`r format(pov_line_latest$pov_line[3], big.mark = ",")` per month, the national poverty line for `r format(pov_line_latest$date[1], "%B %Y")`.

BPS constructs the poverty line based on the food and non-food spending. The food spending is the amount of spending equals 2,100 kilocalories (kcal) for a person a day. The non-food spending comprises the amount of spending on housing, education, healthcare, clothes and other basic needs other than food.

Read more about the data [here](https://www.bps.go.id/subject/23/kemiskinan-dan-ketimpangan.html#subjekViewTab1){target="_blank"}.
