---
title: "Poverty"
description: |
  The COVID-19 pandemic has pushed millions into poverty, undoing around three years worth of progress in lifting people out of it. The rise in the poverty rate was higher in urban areas.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
preview: fig/ier_poverty-rate_void_plot.png
---

```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(htmltools)
library(crosstalk)
library(govdown)
library(mime)

# functions to describe data
## rise or fall
ier_compare_rise <- function(x) {
  if (x > 0) {
    "rose"
  } else {
    "fell"
  }
}

## up or down
ier_compare_up <- function(x) {
  if (x > 0 ) {
    "up"
  } else {
    "down"
  }
}

# embed file
## poverty rate
embed_pov_rate <- xfun::embed_file(
  "data/ier_poverty-rate_cleaned.csv",
  name = "ier_poverty-rate_cleaned.csv",
  text = "csv"
)

## poverty by province
embed_pov_prov <- xfun::embed_file(
  "data/ier_poverty-province_cleaned.csv",
  name = "ier_poverty-province_cleaned.csv",
  text = "csv"
)

# embed chart
## poverty rate
embed_chart_pov_rate <- xfun::embed_file(
  "fig/ier_poverty-rate_plot.png",
  name = "ier_poverty-rate_plot.png",
  text = "chart"
)
```


### Poverty rate

(percent)

```{r plot: poverty rate}
source("script/poverty_script.R", local = knitr::knit_global())
div(plot_pov_rate)
```

*Tip: click the legend item(s) to select or remove data*

```{r description: poverty rate}
# read number of poor people data
source("script/poverty_poor_script.R", local = knitr::knit_global())

# latest number of poor people data
poor_latest <- poor_trf %>% 
  dplyr::filter(date == last(date))

# latest poverty rate data
pov_latest <- pov_trf %>% 
  dplyr::filter(date == last(date))
```

The overall poverty rate `r ier_compare_rise(pov_latest$diff[3])` by **`r pov_latest$diff[3]`** percentage points year-on-year (yoy) to **`r pov_latest$pov_rate[3]`** percent in `r format(pov_latest$date[1], "%B %Y")`. This means the total number of poor people was `r ier_compare_up(poor_latest$diff[3])` by **`r poor_latest$diff[3]`** million yoy to **`r poor_latest$poor[3]`** million.

<aside>

See also [unemployment](employment.html)

</aside>

In urban areas, the poverty rate `r ier_compare_rise(pov_latest$diff[1])` by **`r pov_latest$diff[1]`** percentage points yoy to **`r pov_latest$pov_rate[1]`** percent. In rural areas, the poverty rate was `r ier_compare_up(pov_latest$diff[2])` by **`r pov_latest$diff[2]`** percentage points yoy to **`r pov_latest$pov_rate[2]`** percent.

In September 2020, the pandemic pushed the overall poverty rate back to a level last seen in 2017.

```{r documentation: poverty}
# date of last update
## request data
pressrelease_req <- GET(
  base_url,
  query = list(
    model = "pressrelease",
    lang = "ind",
    domain = "0000",
    keyword = "miskin",
    key = BPS_KEY
  )
)

## parse response
pressrelease_resp <- content(pressrelease_req, "text") %>% 
  fromJSON()

## date of last update
source_last_update <- as_tibble(pressrelease_resp$data[[2]]) %>% 
  dplyr::filter(subj_id == 23) %>% 
  mutate(rl_date = ymd(rl_date)) %>% 
  arrange(desc(rl_date)) %>% 
  dplyr::filter(rl_date == first(rl_date)) %>% 
  select(rl_date) %>% 
  deframe() %>% 
  format("%B %d, %Y")
```

<details>
<summary>Documentation and download options</summary>
**Description**: the share of people who cannot afford both food and non-food basic needs as measured by monthly spending. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every February and July  
**Source**: [Statistics Indonesia (BPS)](https://www.bps.go.id/indicator/23/184/1/persentase-penduduk-miskin-menurut-wilayah.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_pov_rate`; `r embed_chart_pov_rate`  
</details>


#### By province

```{r script: poverty by province}
source("script/poverty_province_script.R", local = knitr::knit_global())

# filter
filter_pov_poor_prov <- filter_select(
  "pov_poor_prov",
  "Select a province...",
  pov_poor_sub_wide_shared,
  ~prov
)

## button
unstrap(filter_pov_poor_prov)
```

```{r table: poverty by province, layout="l-body-outset"}
reactable_pov_poor
```

```{r description: poverty by province}
pov_poor_head <- pov_poor_sub_wide %>% 
  arrange(desc(pov_rate_diff)) %>% 
  head(5)
```

The highest increase in poverty rate was seen in `r pov_poor_head$prov[1]`, where it `r ier_compare_rise(pov_poor_head$pov_rate_diff[1])` by **`r pov_poor_head$pov_rate_diff[1]`** percentage points yoy, which brought the share of poor people in the province to **`r pov_poor_head$pov_rate[1]`** percent in `r format(pov_latest$date[1], "%B %Y")`. It was followed by `r pov_poor_head$prov[-c(1, 5)]` and `r pov_poor_head$prov[5]`.

<aside>

See also [unemployment by province](employment.html#by-province)

</aside>

<details>
<summary>Documentation and download options</summary>
**Description**: the absolute number of people who cannot afford both food and non-food basic needs as measured by monthly spending and the poverty rate, broken down by province. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Update frequency**: every February and July  
**Source**: [BPS](https://www.bps.go.id/indicator/23/192/1/persentase-penduduk-miskin-menurut-provinsi.html){target="_blank"}  
**Last updated**: `r source_last_update`  
**Download**: `r embed_pov_prov`  
</details>


# Note {.appendix}

The data is based on the National Social and Economic Survey (Susenas) from BPS, which takes a sample of roughly 300,000 households for the March round and 75,000 households for the September round.

BPS uses the basic needs approach, which defines poverty as the inability to afford both food and non-food basic needs as measured by spending.

```{r description: poverty line}
source("script/poverty_line.R", local = knitr::knit_global())

# latest poverty line data
pov_line_latest <- pov_line_trf %>% 
  dplyr::filter(date == last(date))
```

Poor people were those who spent below Rp `r format(pov_line_latest$pov_line[3], big.mark = ",")` per month, the national poverty line for `r format(pov_line_latest$date[1], "%B %Y")`.

BPS constructs the poverty line based on the food and non-food spending. The food spending is the amount of spending equals 2,100 kilocalories (kcal) for a person a day. The non-food spending comprises the amount of spending on housing, education, healthcare, clothes and other basic needs other than food.

Read more about the data [here](https://www.bps.go.id/subject/23/kemiskinan-dan-ketimpangan.html#subjekViewTab1){target="_blank"}.
