---
title: "Retail sales"
description: |
  Retail sales have been declining and unusually low during Ramadan, Idul Fitri, Christmas and year end due to the COVID-19 pandemic. While retail sales across categories remain below the usual level before the pandemic, some categories have shown signs of improvements.
output: 
  distill::distill_article:
    toc: true
    toc_float: true
preview: fig/ier_rsi_void_plot.png
---


```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse)
library(xaringanExtra)
library(htmltools)
library(mime)

# panelset
xaringanExtra::use_panelset()

# source url
source_url <- "https://www.bi.go.id/id/publikasi/laporan/Pages/SPE_Januari_2021.aspx"

# date of last update
source_last_update <- "March 9, 2021"

# description functions
## rise or fall
ier_compare_rise <- function(x) {
  if (x > 0) {
    "rise"
  } else {
    "fall"
  }
}

## up or down
ier_compare_up <- function(x) {
  if (x > 0 ) {
    "up"
  } else {
    "down"
  }
}

## less or more severe
ier_compare_sev <- function(x, y) {
  if (x > y) {
    "less severe"
  } else {
    "more severe"
  }
}

# embed file
## overal rsi
embed_rsi_overall <- xfun::embed_file(
  "data/ier_rsi-overall_cleaned.csv",
  name = "ier_rsi-overall_cleaned.csv",
  text = "csv"
)

## rsi by category
embed_rsi_cat <- xfun::embed_file(
  "data/ier_rsi-category_cleaned.csv",
  name = "ier_rsi-category_cleaned.csv",
  text = "csv"
)

# embed chart
## annual change in overall rsi
embed_chart_rsi_overall_chg <- xfun::embed_file(
  "fig/ier_rsi-change_plot.png",
  name = "ier_rsi-change_plot.png",
  text = "chart"
)

## overall rsi
embed_chart_rsi_overall <- xfun::embed_file(
  "fig/ier_rsi_plot.png",
  name = "ier_rsi_plot.png",
  text = "chart"
)

## annual change in rsi by category
embed_chart_rsi_cat_chg <- xfun::embed_file(
  "fig/ier_rsi-cat-change_plot.png",
  name = "ier_rsi-cat-change_plot.png",
  text = "chart"
)

## rsi by category
embed_chart_rsi_cat <- xfun::embed_file(
  "fig/ier_rsi-cat_plot.png",
  name = "ier_rsi-cat-change_plot.png",
  text = "chart"
)
```


### Overall

::::: {.panelset}

::: {.panel}

#### Change {.panel-name}

(percent change from a year earlier)

```{r plot: rsi change}
source("script/rsi_script.R", local = knitr::knit_global())

div(plot_rsi_change)
```

:::

::: {.panel}

#### Index {.panel-name}

```{r plot: rsi}
div(plot_rsi_index)
```

:::

:::::

<p style="color: #999; font-size: 12px;">Note: `r format(last(rsi_trf$date), "%B %Y")` figure is an estimate.</p>  

The Retail Sales Index was estimated to `r ier_compare_rise(last(rsi_trf$pct_change_yoy))` by **`r last(rsi_trf$pct_change_yoy)`** percent year-on-year (yoy) to **`r last(rsi_trf$rsi)`** in `r format(last(rsi_trf$date), "%B")`, `r ier_compare_sev(last(rsi_trf$pct_change_yoy),  rsi_trf$pct_change_yoy[nrow(rsi_trf) - 1])` than the previous month. 

<aside>

See also [community mobility](mobility.html)

</aside>

The pandemic has led the index to fall to its steepest decline at **20.61** percent yoy to **198.3** in May 2020, when Ramadan and Idul Fitri took place.

<details>
<summary>Documentation</summary>
**Description**: the Retail Sales Index indicates the retail sales performance in a given month. The base year for the index is January 2010. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Source**: [Bank Indonesia (BI)](`r source_url`){target="_blank"}  
**Data and update frequency**: monthly  
**Last updated**: `r source_last_update`  
**Download**: `r embed_rsi_overall`; index (`r embed_chart_rsi_overall`); change (`r embed_chart_rsi_overall_chg`)  
</details>


### By category

::::: {.panelset}

::: {.panel}

#### Change {.panel-name}

(percent change from a year earlier)

```{r plot: rsi change by category}
source("script/rsi_category_script.R", local = knitr::knit_global())

div(sm_rsi_by_cat_chg)
```

:::

::: {.panel}

#### Index {.panel-name}

```{r plot: rsi by category}
div(sm_rsi_by_cat_idx)
```

:::

:::::

```{r description: rsi by category}
# categories showing improvements
rsi_by_cat_improve <- rsi_by_cat_trf %>% 
  group_by(Indices) %>% 
  dplyr::filter(last(pct_change_yoy) > tail(pct_change_yoy, 2)[[1]]) %>% 
  ungroup() %>% 
  select(Indices) %>% 
  dplyr::filter(!duplicated(Indices))

# set to lower case
rsi_by_cat_imp_lower <- str_to_lower(rsi_by_cat_improve[[1]])

# categories showing steeper declines
rsi_by_cat_worse <- rsi_by_cat_trf %>% 
  group_by(Indices) %>% 
  dplyr::filter(
    last(pct_change_yoy) < tail(pct_change_yoy, 2)[[1]],
    date == last(date)
  )

# set to lower case
rsi_by_cat_worse_lower <- str_to_lower(rsi_by_cat_worse$Indices)
```

<div style="color: #999; font-size: 12px;">
<sup>*</sup>Information and communication tools  
<sup>&#10013;</sup>Cultural and recreational goods  
Note: `r format(last(rsi_trf$date), "%B %Y")` figure is an estimate.  
</div>

The Retail Sales Index of `r rsi_by_cat_imp_lower[-length(rsi_by_cat_imp_lower)]` and `r rsi_by_cat_imp_lower[length(rsi_by_cat_imp_lower)]` improved in `r format(last(rsi_trf$date), "%B")` from the previous month in terms of change from a year earlier.

<aside>

See also [inflation by expenditure group](inflation.html#by-expenditure-group)

</aside>

Meanwhile, `r rsi_by_cat_worse_lower[-length(rsi_by_cat_worse_lower)]` and `r rsi_by_cat_worse_lower[length(rsi_by_cat_worse_lower)]` showed steeper declines than the previous month in terms of percent change from a year earlier.

<details>
<summary>Documentation</summary>
**Description**: the Retail Sales Index, broken down by category, also indicates the retail sales performance in a given month. The base year for the index is January 2010. See [appendix](#note) for further details.  
**Chart**: Dzulfiqar Fathur Rahman  
**Source**: [BI](`r source_url`){target="_blank"}  
**Data and update frequency**: monthly  
**Last updated**: `r source_last_update`  
**Download**: `r embed_rsi_cat`; index (`r embed_chart_rsi_cat`); change (`r embed_chart_rsi_cat_chg`)  
</details>


# Note {.appendix}

The Retail Sales Index is based on a monthly survey from Bank Indonesia. The central bank has conducted the survey since September 1999.

Since January 2015, the survey covers around 700 retailers in 10 cities, comprising Jakarta, Semarang, Bandung, Surabaya, Medan, Purwokerto, Makassar, Manado, Banjarmasin and Denpasar.